Species turnover reveals hidden effects of decreasing Nitrogen deposition in mountain hay meadows

### 2018-10-24 Manuscript.Rmd
Discuss rather large pseudo-turnover.

For each survey (that consisted of two visits per season) we calculated the following biodiversity endpoints [@Rowe2017]: We used the number of recorded species (species richness) that can be easily related to many conservation targets [@Rowe2017]. Additionally, we calculated the community mean of the Landolt indicator values of recorded species.Similar to the Ellenberg’s indicator values [@Ellenberg1974], the Landolt values are ordinal numbers that express the realized ecological optima of plants species for different climatic, soil or land-use variables. The Landolt indicator values were developed for the specific situation in Switzerland, published the first time in @Landolt1977 and recalibrarted in @Landolt2010. Their predictive power was tested in different studies [e.g. @Scherrer2011].We analysed the indicator values for temperature (1: high elevation species; 5: low elevation species), soil moisture (1: species that grow in soils with low moisture; 5: species that grow in water-saturated soils), nutrients (in particular nitrogen, but also phosphorus; 1: species that grow in nutrient-poor soils; 5: species that grow under nutrient-rich conditions) and light (1: species that grow in shade; 5: species that predominantly occur in bright places).

@@

## Plant communities {-}
In total, `r n_distinct(pl$aID_SP)` plant species were recorded on the `r nrow(sites)` plots. Including the data of all three visits, `r mean(surv$SR)` $\pm$ `r sd(surv$SR)` (average $\pm$ SD) species were observed per plot. The lowest number of species recorded during a survey was `r min(surv$SR)` species and the highest number was `r max(surv$SR)` species. The average indicator value for temperature accross all surveys was `r mean(surv$T)` $\pm$ `r sd(surv$T)` ranging from `r min(surv$T)` to `r max(surv$T)`. The average indicator value for soil moisture was `r mean(surv$F)` $\pm$ `r sd(surv$F)` ranging from `r min(surv$F)` to `r max(surv$F)`. The average indicator value for nutrients was `r mean(surv$N)` $\pm$ `r sd(surv$N)` ranging from `r min(surv$N)` to `r max(surv$N)`. And the average indicator value for light was `r mean(surv$L)` $\pm$ `r sd(surv$L)` ranging from `r min(surv$L)` to `r max(surv$L)`.

@@

# Compile res for table
res <- tibble()
tparams <- c("Temperature", "Precipitation", "NTOT", "Inclination")
res <- tibble(
  Gradient = c("Mean annual temperature", "Mean annual precipitation", "Nitrogen deposition", "Inclination"),
  Estimate = fixef(mod)[tparams],
  `5%` = posterior_interval(mod, pars = tparams)[,1],
  `95%` = posterior_interval(mod, pars = tparams)[,2])

# 90% CrI for period effect
tmp <-  posterior_interval(mod, pars = "period")
CrI <- paste0(round(tmp[1,1], 2), " - ", round(tmp[1,2], 2))
```

```{r turnovertab, echo=FALSE}
options(knitr.kable.NA = '')
kable(res, align = c("l", rep("r", 3)), digits = 2,  booktabs = T, col.names = c("Gradient", "Estimate", "5%", "95%"), caption = "Change of species turnover along the four gradients. The slopes along the gradients (estimate) are given as the change per 10 years of the logit-probability of species that differed between two surveys. Estimates and 5\\% and 95\\% quantiles of the marginal posterior distribution obtained from a binomial GLMM.")
```

The average $\pm$ SD percentage of species that differed between replicated surveys (i.e. different botanists surveyed the sites) was `r mean(100 * turnover_double$turnover)` $\pm$ `r sd(100 * turnover_double$turnover)`%. This turnover between replicated surveys was lower than the observed temporal turnover: the average percentage of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`%, and the percentage of species that differed between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%. Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (90% credible interval of the change in turnover estimated from the binomial GLMM: `r CrI`). Along the N deposition gradient, variation in species turnover was largest, with highest species turnover at sites with low N deposition (Table \@ref(tab:turnovertab)). The other three gradients were less conclusive in explaining the variation in species turnover among sites.

@@

>>>>

Before the analyses we removed all records that were not identified on species level. For each survey (that consisted of two visits per season) we then calculated the following biodiversity endpoints [@Rowe2017]: We used the number of recorded species (species richness) that can be easily related to many conservation targets [@Rowe2017]. Additionally, we calculated the community mean of the Landolt indicator values of recorded species.Similar to the Ellenberg’s indicator values [@Ellenberg1974], the Landolt values are ordinal numbers that express the realized ecological optima of plants species for different climatic, soil or land-use variables. The Landolt indicator values were developed for the specific situation in Switzerland, published the first time in @Landolt1977 and recalibrarted in @Landolt2010. Their predictive power was tested in different studies [e.g. @Scherrer2011].We analysed the indicator values for temperature (1: high elevation species; 5: low elevation species), soil moisture (1: species that grow in soils with low moisture; 5: species that grow in water-saturated soils), nutrients (in particular nitrogen, but also phosphorus; 1: species that grow in nutrient-poor soils; 5: species that grow under nutrient-rich conditions) and light (1: species that grow in shade; 5: species that predominantly occur in bright places).

@@

## Plant communities {-}
In total, `r n_distinct(pl$aID_SP)` plant species were recorded on the `r nrow(sites)` plots. Including the data of all three visits, `r mean(surv$SR)` $\pm$ `r sd(surv$SR)` (average $\pm$ SD) species were observed per plot. The lowest number of species recorded during a survey was `r min(surv$SR)` species and the highest number was `r max(surv$SR)` species. In average `r mean(100 * surv$KL)` $\pm$ `r sd(100 * surv$KL)`% of the recorded species were annual species with one plot reaching up to `r mean(100 * surv$KL)`% annual species. The average indicator value for temperature accross all surveys was `r mean(surv$T)` $\pm$ `r sd(surv$T)` ranging from `r min(surv$T)` to `r max(surv$T)`. The average indicator value for soil moisture was `r mean(surv$F)` $\pm$ `r sd(surv$F)` ranging from `r min(surv$F)` to `r max(surv$F)`. The average indicator value for nutrients was `r mean(surv$N)` $\pm$ `r sd(surv$N)` ranging from `r min(surv$N)` to `r max(surv$N)`. And the average indicator value for light was `r mean(surv$L)` $\pm$ `r sd(surv$L)` ranging from `r min(surv$L)` to `r max(surv$L)`.

@@

# Compile res for table
res <- tibble()
tparams <- c("period", "SR", "Temperature", "Precipitation", "NTOT", "Inclination")
res <- tibble(
  Predictors = c("Period", "Number of species", "Mean annual temperature", "Mean annual precipitation", "Nitrogen deposition", "Inclination"),
  Estimate = fixef(mod)[tparams],
  `5%` = posterior_interval(mod, pars = tparams)[,1],
  `95%` = posterior_interval(mod, pars = tparams)[,2])
```

```{r turnovertab, echo=FALSE}
options(knitr.kable.NA = '')
kable(res, align = c("l", rep("r", 3)), digits = 2,  booktabs = T, col.names = c("Predictors", "Estimate", "5%", "95%"), caption = "Change of species turnover along the four gradients when differences between the two periods (first period: turnover between first and second survey; second period: change in turnover between second and third surveys) and species richness effects are acounted for. Estimates for the period effect (change in turnover from first to second survey), the species richness effect (change in turnover per 10 species) and along the four gradients (slopes) with the corresponding 5\\% and 95\\% quantiles of the marginal posterior distribution were obtained from a binomial GLMM.")
```

The average $\pm$ SD percentage of species that differed between replicated surveys (i.e. different botanists surveyed the sites) was `r mean(100 * turnover_double$turnover)` $\pm$ `r sd(100 * turnover_double$turnover)`%. This turnover between replicated surveys was lower than the observed temporal turnover: the average percentage of species that differed between the first and second survey at a site was `r mean(100 * sites$change1 / (sites$nochange1 + sites$change1))` $\pm$ `r sd(100 * sites$change1 / (sites$nochange1 + sites$change1))`%, and the percentage of species that differed between the second and third survey was `r mean(100 * sites$change2 / (sites$nochange2 + sites$change2))` $\pm$ `r sd(100 * sites$change2 / (sites$nochange2 + sites$change2))`%. Thus, it seemed that the turnover from the first/second survey to the turnover of the second/third survey moderately decreased (period effect in Table \@ref(tab:turnovertab)). Species richness was a good predictor of species turnover: species rich sites were subject to higher turnover than sites with lower species richness (Table \@ref(tab:turnovertab)). The four gradients (temperature, precipitation, Nitrogen deposition and inclination) were less conclusive in explaining the variation in species turnover among sites.

@@

Our results conform to the results of recent reviews on biodiversity change suggesting that local-scale species communities are often undergoing profound changes but not systematic loss [@Dornelas2014]. However, our comparision with replicated surveys warn that an important portion of the observed turnover of species might be due to pseudo-turnover (i.e. species difference between two surveys that were conducted by two surveyors during the same year on the same site). Given that the BDM program had invested significantly in developing reproducable methods and has continiously invested in quality control [@Plattner2004], the recorded pseudo-turnover was quite high. A potential explanation is that the replicated surveys are not conducted during the same days and in few cases the situation might have changeed profoundly for example because the meadow was cut between the surveys. Furthermore, one of the botanist may not have identified difficult to identify species on species-level, and because we analysed only records that were identified on species level this increased pseudo-turnover even tough both botanists detected the species.

- Trotzdem finden wir Muster, die kaum auf Methodische Gründe Zurückzuführen sind. Zum Beispiel ist Turnover bei tiefer N Dep am höchsten. Aber warum ist das so?
- Ev auch noch den Anteil der Perennial diskutieren --> Turnover ist erstaunlich hoch.
